---
title: "Fuel economy Presentation"
author: "Odia"
date: "2026-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

This report presents an exploratory data analysis (EDA) of the `mpg` dataset. The dataset contains fuel economy information for various car models manufactured between 1999 and 2008. The objective is to understand the relationships between vehicle characteristics and fuel efficiency.

# Dataset Description

The `mpg` dataset contains **234 observations** and **11 variables**.

| Variable | Description |
|----------|-------------|
| manufacturer | Car manufacturer |
| model | Model name |
| displ | Engine displacement (litres) |
| year | Model year |
| cyl | Number of cylinders |
| trans | Transmission type |
| drv | Drive type (front, rear, four-wheel) |
| cty | City miles per gallon |
| hwy | Highway miles per gallon |
| fl | Fuel type |
| class | Vehicle class |

```{r include=FALSE}
library(tidyverse)
```

```{r description, echo=FALSE}
glimpse(mpg)
summary(mpg)
```

# Data Exploration

## First Six Observations

```{r first 5 rows, echo=FALSE}
head(mpg)
```

## Structure of the Dataset

```{r echo=FALSE}
str(mpg)
```

## Summary Statistics

```{r echo=FALSE}
summary(mpg)
```

# Data Visualization

## Distribution of Highway Mileage

```{r echo=FALSE}
ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Highway MPG",
    x = "Highway MPG",
    y = "Frequency"
  )
```

## Engine Size vs Highway MPG

```{r echo=FALSE}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Engine Displacement vs Highway MPG",
    x = "Engine Displacement (L)",
    y = "Highway MPG"
  )
```

## Average Highway MPG by Vehicle Class

```{r echo=FALSE}
mpg %>%
  group_by(class) %>%
  summarise(avg_hwy = mean(hwy)) %>%
  ggplot(aes(reorder(class, avg_hwy), avg_hwy)) +
  geom_col(fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Average Highway MPG by Vehicle Class",
    x = "Vehicle Class",
    y = "Average Highway MPG"
  )
```

# Key Findings

- The dataset contains information on **234 vehicles** from multiple manufacturers.
- Smaller engines generally achieve better highway fuel economy.
- Compact and subcompact vehicles tend to have the highest average highway MPG.
- SUVs and pickup trucks generally have lower fuel efficiency.
- There is a negative relationship between engine displacement and fuel economy.

# Conclusion

The exploratory analysis shows that vehicle characteristics such as engine displacement and vehicle class significantly influence fuel efficiency. Cars with smaller engines generally consume less fuel and achieve higher highway mileage, while larger vehicles tend to be less fuel efficient. Further analysis could investigate how transmission type or drive type affects fuel economy.